---
title: "Untitled"
author: "Wayne Zhang"
date: '2022-03-09'
output: html_document
---

# Basic scraping

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

```{r}
link <- "https://www.imdb.com/search/title/?title_type=feature&my_ratings=exclude&adult=include&num_votes=250000,"

page <- read_html(link)
```

The information we need here is name, year, rating and the abstract.

```{r}
name <- page %>% 
    html_nodes(".lister-item-header a") %>% 
    html_text() # from SelectorGadget to get the nodes name
year <- page %>% 
    html_nodes(".text-muted.unbold") %>% 
    html_text() %>% 
    str_sub(-5, -2) %>% 
    as.numeric()
rating <- page %>% 
    html_nodes(".ratings-imdb-rating strong") %>% 
    html_text() %>% 
    as.numeric()
abstract <- page %>% 
    html_nodes(".ratings-bar+ .text-muted") %>% 
    html_text() %>% 
    str_remove("\\n")

imdb <- data.frame(year, name, rating, abstract) %>% 
    arrange(desc(rating))

knitr::kable(head(imdb))
```

## Scraping from the links

The next step is we want the cast list of each movie. We ca get it from after click the movie name and it was contained in the Top cast.

```{r}
# the first step is to get the url from the name of movie
# by observe the url of the page, we know that the url is "https://www.imdb.com" + herf
movie_link <- page %>% 
    html_nodes(".lister-item-header a") %>% # this part is same as name
    html_attr("href") %>% 
    str_replace("\\?ref_=adv_li_tt", "fullcredits?ref_=tt_cl_sm") %>% 
    paste("https://www.imdb.com", ., sep = "")

# we can not pipe read_html_again, because currently movie_link is a vector
get_cast <- function(movie_link){
    movie_html <- read_html(movie_link)
    cast <- movie_html %>% 
        html_nodes(".primary_photo+ td a") %>% 
        html_text() %>% 
        str_remove("\\n")
    name <- movie_html %>% 
        html_nodes(".parent a") %>% 
        html_text()
    data.frame(name = rep(name, length(cast)), cast)
}

cast <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(cast) <- c("name", "cast")
for(i in movie_link){
    cast <- rbind(cast, get_cast(i))
}

knitr::kable(cast %>% 
                 group_by(name) %>% 
                 summarise(first_cast = first(cast)))
```

## Scraping from multiple pages

The next step we want to do is to scrapping all movies (previous we only scrap the first 50 movies)

```{r}
# new link: https://www.imdb.com/search/title/?title_type=feature&num_votes=250000,&my_ratings=exclude&adult=include&start=51&ref_=adv_nxt
# previous link: https://www.imdb.com/search/title/?title_type=feature&my_ratings=exclude&adult=include&num_votes=250000,
imdb2 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(imdb2) <- c("name", "year", "rating")
for(i in seq(1, 751, 50)){
    link <- paste("https://www.imdb.com/search/title/?title_type=feature&num_votes=250000,&my_ratings=exclude&adult=include&start=",i , "&ref_=adv_nxt", sep = "")
    page <- read_html(link)
    name <- page %>% 
        html_nodes(".lister-item-header a") %>% 
        html_text() # from SelectorGadget to get the nodes name
    year <- page %>% 
        html_nodes(".text-muted.unbold") %>% 
        html_text() %>% 
        str_sub(-5, -2) %>% 
        as.numeric()
    rating <- page %>% 
        html_nodes(".ratings-imdb-rating strong") %>% 
        html_text() %>% 
        as.numeric()
    imdb2 <- rbind(imdb2, data.frame(year, name, rating))
}

knitr::kable(tail(imdb2, 20))
```

## Scraping table

```{r}
link <- "https://www.patriotsoftware.com/blog/accounting/average-cost-living-by-state/"
page <- read_html(link)

price_chart <- page %>% 
    html_nodes("table") %>% 
    html_table()

knitr::kable(head(price_chart))
```

```{r}
# another example
wiki_link <- "https://en.wikipedia.org/wiki/List_of_apple_cultivars"
wiki_page <- read_html(wiki_link)

# Table of apples
table_apple <- wiki_page %>% 
    html_nodes("table#wikitable.sortable.jquery-tablesorter") %>% 
    html_table() # here it return a empty list
# the reason is because the css framework, wiki_page is different than we see in wiki page
# the solution is to check the tag from wiki_page directly

wiki_page %>% 
    html_nodes("table")
 
# and we know the second one is what we want
 
table_apple <- wiki_page %>% 
    html_nodes("table") %>% 
    .[2] %>% 
    html_table() %>% 
    .[[1]] %>% # we only use its first element
    select(-Image)
knitr::kable(head(table_apple))
```


# Application

Scrap the data from OECD

```{r}
# link <- "https://data.oecd.org/price/inflation-cpi.htm"
# page <- read_html(link)
# 
# page %>%
#     html_nodes(".indicator-group-navigation-list :nth-child(1)") %>%
#     html_attr("data-indicator")
# data <- read.csv("https://stats.oecd.org/sdmx-json/data/DP_LIVE/.CPIFORECAST.../OECD?contentType=csv&detail=code&separator=comma&csv-lang=en")
```

