---
title: "Time series review"
author: "Wayne Zhang"
date: '2022-03-29'
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(FinCal)
library(Statamarkdown)
```

# Core Model

* Autoregressive (AR)

$$y_t = \alpha + \phi_1 y_{t - 1} + \cdots + \phi_p y_{t - p} + e_t$$

* Regression 

$$y_t = \alpha + \delta_0 x_t + e_t$$

* Distributed Lag (DL)

$$y_t = \alpha + \delta_0 x_t + \delta_1 x_{t - 1} + \cdots + \delta_q x_{t - q} + e_t$$

* Autoregressive-Distributed Lag(ADL)

$$y_t = \alpha + \phi_1 y_{t - 1} + \cdots + \phi_p y_{t - p} + \delta_0 x_t + \delta_1x_{t - 1}+\cdots + \delta_q x_{t - q} + e_t$$

**Remark**: The coefficients in the DL and ADL can be interpreted as multipliers. Multipliers are structural under exogeneity;

## Standard Errors:

Three methods:

1. Classical (homoskedastic): **not used** in contemporary economics.
2. Robust (Heteroskedastic): 
    * Appropriate for AR and ADL;
    * **Inappropriate** for  non-dynamic regression or DL.
3. HAC (Newey-West):
    * Important for regression or DL.

**Remark**:

* **Always** report coefficient estimates & standard errors; Reason: Standard errors convey degree of uncertainty -- always important
* **Never** coefficient estimates & t-statistics; t-statistics concern testing hypothesis of a zero coefficient -- rarely of key interest

# Application:

## AR model

Illustration: U.S. quarterly real GDP growth rates, post-war

```{r ar_model1}
gdp <- read_dta("/Users/waynezhang/Library/CloudStorage/OneDrive-UW-Madison/IOS-WIN/Econometric method/EMLearn/Software/R/study_R/time_series_21/gdp.dta")

stata("use gdp\n reg gdp L(1/3).gdp, r", data.in = gdp)

library(lmtest)
library(sandwich)
sample_gdp <- gdp$gdp
mod_ar <- lm(sample_gdp ~ lag(sample_gdp, 1) + lag(sample_gdp, 2) + lag(sample_gdp, 3))
coeftest(mod_ar, vcov. = vcovHC, type = "HC1")
```

```{r ar_model2}
prices <- stata("use prices", data.out = T)

stata("use prices\n gen return = 100*(sp500/L.sp500-1)\n reg return L(1/2).return ,r\n testparm L(1/2).return", data.in = prices)

return <- (prices$sp500/lag(prices$sp500) - 1)*100
mod_ar2 <- lm(return ~ lag(return, 1) + lag(return, 2))
coeftest(mod_ar2, vcov. = vcovHC, type = "HC1")
```


## DL model

Illustration: retail gasoline and crude oil prices


```{r}
stata("use prices\n 
        gen dg = 100*(gasoline/L.gasoline - 1)\n 
        gen db = 100*(brent/L.brent - 1) \n
        reg dg db, r\n
        newey dg db, lag(12)\n
        newey dg L(0/6).db, lag(12)\n
        newey dg L(0/5).d.db L6.db, lag(12)
      ")
# lag: set maximum lag order of autocorrelation
dg = 100*(prices$gasoline/lag(prices$gasoline) - 1)
db = 100*(prices$brent/lag(prices$brent) - 1)
dat <- as.data.frame(na.omit(cbind(dg, db)))
dg <- dat$dg
db <- dat$db

mod_dl <- lm(dg ~ db)
coeftest(mod_dl, vcov. = vcovHC, type = "HC1")

mod_dl2 <- lm(dg ~ db)
coeftest(mod_dl, vcov. = NeweyWest(mod_dl2, lag = 12))

dat <- dat %>% 
    mutate(lag_db1 = lag(db),
           lag_db2 = lag(lag_db1),
           lag_db3 = lag(lag_db2),
           lag_db4 = lag(lag_db3),
           lag_db5 = lag(lag_db4),
           lag_db6 = lag(lag_db5))
formula <- "dg ~ db"
for (i in 1:6){
    formula <- paste0(formula, " + lag_db", i)
}
mod_dl2 <- lm(formula, dat)
coeftest(mod_dl2, vcov. = NeweyWest(mod_dl2, lag = 12))
```


## ADL

```{r}
stata("use gdp\n 
        gen inf = 400*(cpi/L.cpi-1)\n
        gen dinf = d.inf\n
        reg dinf L(1/4).dinf L(1/4).ur if time<=tq(2016q3), r")
gdp <- stata("use gdp", data.out = T)
gdp <- gdp %>% 
    mutate(inf = 400*(cpi/lag(cpi)-1),
           dinf = inf - lag(inf)) %>% 
    filter(time < 227)

formula <- "dinf ~ "
for (i in 1:4){
    formula <- paste0(formula, "lag_dinf", i, " + ")
    gdp[, ncol(gdp) + 1] <- lag(gdp$dinf, i)
    names(gdp)[ncol(gdp)] <- paste0("lag_dinf", i) 
}
formula <- substr(formula,1,nchar(formula)-3)
for (i in 1:4){
    formula <- paste0(formula, " + lag_ur", i)
    gdp[, ncol(gdp) + 1] <- lag(gdp$ur, i)
    names(gdp)[ncol(gdp)] <- paste0("lag_ur", i) 
}
mod_adl <- lm(formula, gdp)
coeftest(mod_adl, vcov. = vcovHC, type = "HC1")
```

