---
title: "Causal Inference6"
author: "Weiheng Zhang"
date: '2022-04-22'
output: 
    pdf_document:
        extra_dependencies: ["amsfonts"]
---

# Regression Discountinuity Designs

## Sharp RD

This is a sharp RD because the probability of treatment switches from zero zero to one at the threshold.

\begin{itemize}
\item Basic Setup

Suppose the treatment is a deterministic and discontinuous function of observed covariate $R_i$ such that
    
$$ D_i = 1\{X_i > c\}$$

where $X_i$ is called the running variable or forcing variable. c is determined by policy or rule. An individual is untreated if D = 0 and is treated if D = 1.




\item Identification

The random individual treatment effect $\theta(x) = E[\theta|X = x]$ is the conditional ATE. $\bar\theta = \theta(c)$, the conditional ATE for the subpopulation at the cut-off.

$$\begin{aligned}
m(x) &= E[Y|X = x]\\
m_0(x) &= E[Y_0|X = x]\\
m_1(x) &= E[Y_1|X = x] \\
\theta(x) &= m_1(x) - m_0(x)
\end{aligned}$$

\end{itemize}

### Application Carpenter and Dobkin (2009)

```{r}
data(mm_mlda, package = "stevedata")
# Scatterplot
mm_mlda <- mm_mlda %>% 
    drop_na()
mm_mlda %>% 
    ggplot(aes(x = agecell, y = all)) +
    geom_point() +
    geom_vline(xintercept = 21, color = "red", linetype = "dashed") +
    annotate("text", x = 20.4, y = 105, label = "Minimum Drinking Age") +
    labs(y = "Mortality rate (per 100.000)",
         x = "Age (binned)")
```

```{r}
# Estimate the model (same slopes)

# Create threshold variable
mm_mlda$threshold <- ifelse(mm_mlda$agecell >= 21, 1, 0)
lm_same_slope <- mm_mlda %>% 
    lm(all ~ threshold + I(agecell - 21), .)
summary(lm_same_slope)

# Scatterplot
mm_mlda %>% 
    select(agecell, all, threshold) %>% 
    mutate(fit = predict(lm_same_slope)) %>% 
    ggplot() +
    geom_point(aes(x = agecell, y = all, color = factor(threshold))) +
    geom_line(aes(x = agecell, y = fit, group = threshold)) +
    geom_vline(xintercept = 21, color = "red", linetype = "dashed") +
    guides(color = FALSE) +
    labs(y = "Mortality rate (per 100.000)",
         x = "Age (binned)")
```


```{r}
# Estimate the model (different slopes)

lm_different_slope <- mm_mlda %>% 
    lm(all ~ threshold + I(agecell - 21) + threshold:I(agecell - 21), data = .)
summary(lm_different_slope)

# Scatterplot
mm_mlda %>% 
    select(agecell, all, threshold) %>% 
    mutate(fit = predict(lm_different_slope)) %>% 
    ggplot() +
    geom_point(aes(x = agecell, y = all, color = factor(threshold))) +
    geom_line(aes(x = agecell, y = fit, group = threshold)) +
    geom_vline(xintercept = 21, color = "red", linetype = "dashed") +
    guides(color = FALSE) +
    labs(y = "Mortality rate (per 100.000)",
         x = "Age (binned)")
```

```{r}
# More complicated functional form

lm_quadratic <- mm_mlda %>% 
    lm(all ~ threshold + I(agecell - 21) + I((agecell - 21)^2) + 
           threshold:I(agecell - 21) + threshold:I((agecell - 21)^2), data = .)
summary(lm_quadratic)
# treatment effect 9.5478

# Scatterplot
mm_mlda %>% 
    select(agecell, all, threshold) %>% 
    mutate(fit = predict(lm_quadratic)) %>% 
    ggplot() +
    geom_point(aes(x = agecell, y = all, color = factor(threshold))) +
    geom_line(aes(x = agecell, y = fit, group = threshold)) +
    geom_vline(xintercept = 21, color = "red", linetype = "dashed") +
    guides(color = FALSE) +
    labs(y = "Mortality rate (per 100.000)",
         x = "Age (binned)")
```

```{r}
# sensitivity analysis (local regression)
lm_sensitivity <- mm_mlda %>%
  filter(agecell >= 20 & agecell <= 22) %>%
  mutate(threshold = ifelse(agecell >= 21, 1, 0)) %>% 
  lm(all ~ threshold + I(agecell - 21) + threshold:I(agecell - 21), data = .)

summary(lm_sensitivity)
```


## Fuzzy RD

```{r}
library(MASS)

# generate sample data
mu <- c(0, 0)
sigma <- matrix(c(1, 0.7, 0.7, 1), ncol = 2)

set.seed(1234)
d <- as.data.frame(mvrnorm(2000, mu, sigma))
colnames(d) <- c("W", "Y")

# introduce fuzziness
d$treatProb <- ifelse(d$W < 0, 0, 0.8)

d$fuzz <- sapply(X = d$treatProb, FUN = function(x) rbinom(1, 1, prob = x))

# treatment effect
d$Y <- d$Y + d$fuzz * 2

ggplot(d, aes(x = W, y = Y, color = factor(fuzz))) +
    geom_point() +
    scale_color_manual(values = c("steelblue", "darkred"))
```

