---
title: "Causal Inference7"
author: "Weiheng Zhang"
date: '2022-04-23'
output: 
  pdf_document: 
    toc: yes
---

```{r, message=FALSE, warning=FALSE}
library(FinMetric)
library(formatR)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),
                      tidy=TRUE,
                      echo = TRUE)
```


# Binary Choice

```{r}
# loading dataset
data("Hdma", package = "Ecdat")

Hdma <- Hdma %>% 
    mutate(deny_num = ifelse(deny == "no", 0, 1),
           black_numb = ifelse(black == "no", 0, 1),
           ccs = round(ccs, digits = 0))
```


## Linear Probability Model

$$\begin{aligned}
Y &= P(x) + e\\
E[e|x] &= 0
\end{aligned}$$

```{r}
# Baseline Model
model_lpm1 <- lm(deny_num ~ dir, data = Hdma)
coeftest(model_lpm1, vcov. = vcovHC, type = "HC1")

ggplot(Hdma, aes(dir, deny_num)) +
    geom_point() + 
    geom_abline(slope = model_lpm1$coefficients[2],
                intercept = model_lpm1$coefficients[1]) +
    scale_y_continuous(breaks = c(0,1), limits = c(-0.5,1.5)) +
    geom_hline(yintercept = 1, color = "red", lty = "dashed") +
    geom_hline(yintercept = 0, color = "red", lty = "dashed") +
    labs(x = "Debt to Income Ratio", y = "Mortgage Application Status")
```

The estimate of coefficient `dir` is .604. Doubling the loan payment to income ratio (an incrase by one unit) leads to an increase of the probability of loan denial by 60.4%.

## Index model

$$P[Y_i = 1|X_i] = G(X_i\beta)$$

where $G$ is the link function and $X_i\beta$ is index. Link function could be normal distribution(Probit) or Logistics(Logit).

```{r}
model_logit <- glm(deny_num ~ dir + black,
                   data = Hdma,
                   family = binomial(link = "logit"))

ggplot(data = Hdma,
       aes(x = dir, y = deny_num)) +
  geom_point() +
  stat_smooth(method="glm", method.args=list(family=binomial(link="logit"))) +
  scale_y_continuous(breaks = c(0,1), limits = c(-0.5,1.5)) +
  geom_hline(yintercept = 1, color = "red", lty = "dashed") +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Debt to Income Ratio", y = "Mortgage Application Denied?")

# calculate Prob (Y = 1)
```

$$P(Y_i = 1) = \frac{1}{1 + exp(-(X'\beta + u))}$$

```{r}
value <- c(1, Hdma$dir[1], Hdma$black_numb[1])
1/(1 + exp(-sum(value*coefficients(model_logit))))

fitted(model_logit)[1]


# likelihood ratio test: Similar to an F-test it checks whether all regressors in the model are together equal to zero. 
lrtest(model_logit)
# Pseudo R^2
1 - model_logit$deviance/model_logit$null.deviance
# Log-likelihood
logLik(model_logit)
```

#### Fixed effect in index model

```{r}
model_logit2 <- glm(deny_num ~ dir + factor(ccs),
                    data = Hdma,
                    family = binomial(link = "logit"))

# Predict power
data.frame(dir = mean(Hdma$dir),
           ccs = factor(rep(1:6))) %>% 
    mutate(pred.prob = predict(model_logit2, newdata = .,
                        type = "response"),
           se = predict(model_logit2,newdata = .,
                        type = "response", se = T)$se.fit,
           ll = pred.prob - (1.96*se),
           ul = pred.prob + (1.96*se)) %>% 
    ggplot(data = ., aes(x = ccs, y = pred.prob)) +
    geom_errorbar(aes(ymin = ll, ymax = ul),
                  width = .2, color = "red") +
    geom_point(fill = "black") +
    labs(x = "Consumer's credit score",
         y = "Predicted Probability (of loan denial)")

```


```{r}
# Probit model
model_probit <- glm(deny_num ~ dir + black,
                    data = Hdma,
                    family = binomial(link = "probit"))

ggplot(data = Hdma,
       aes(x = dir, y = deny_num)) +
  geom_point() +
  stat_smooth(method="glm", method.args=list(family=binomial(link="probit")),
              se = F) +
  scale_y_continuous(breaks = c(0,1), limits = c(-0.5,1.5)) +
  geom_hline(yintercept = 1, color = "red", lty = "dashed") +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Debt to Income Ratio", y = "Mortgage Application Status")
```

### Marginal Effects

```{r}
library(margins)
# Average marginal effects (AME)
margins(model_logit2)

ggplot(data = summary(margins(model_logit))) +
  geom_point(aes(factor, AME)) +
  geom_errorbar(aes(x = factor, ymin = lower, ymax = upper), 
                width = 0.2) +
  labs(x = "")
```
















