---
title: "TSRethinking1 Change points and structural breaks"
author: "Wayne Zhang"
date: "10/9/2021"
output:
  html_document: 
    toc: yes
    toc_depth: 4
---
\newcommand{\df}[1]{\textbf{Define #1}}

```{r package, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 4)
rm(list=ls())  # remove variables
library(changepoint) # changing point
library(strucchange) # structural change
library(tidyverse)
library(lubridate)
library(FinCal)
```


Time series object:

- `base`: `ts` object
- `xts` or `zoo`, based on respected packages;
- `tidyverse` object.

## 1 Introduction

Problem: Time series variables may be influenced by events that arise at various points in time that alter the underlying data generating process, which influence the results of many linear time series models.

### 1.1 Change point detection

\df{Changing point CP}: For TS $y_t = \{y_i\}_1^T$ there is a CP, a time, if $\exists \tau\in 1:(T - 1)$ s.t. $\{y_i\}_1^\tau, \{y_i\}_{\tau + 1}^T$ have different properties.

#### Step 1 identify CP

Hypothesis:

$$\begin{aligned} H_0:& \text{ there are no CP}\\ H_1:& \exists \text{ a CP} \end{aligned}$$

method: MLE, likelihood ratio test

#### Step 2 detect the position of CPs

detect multiple CPs: thought minimizing cost function + penalty function ($f(m)$ is a penalty function could take different information criteria, like AIC, BIC, etc.)

$$\sum^{m+1}_{i=1} \left[ \mathcal{C} \left( y_{(\tau_{i-1}+1):\tau_{1}} \right) \right] + \beta f (m)$$

- Binary segmentation (critique): apply single CP method to all data and the if there is a CP using the CP seperate futher subsample applying single CP test again.

- Segment neighborhood: dynamic programming technique to get the optimal segmentation.

- The pruned exact linear time (PELT) algorithm also provides an exact solution, but is more computationally efficient

##### Change point in mean

```{r simulate_mean}
set.seed(123)
sim_mean <- c(rnorm(100, 0, 1),
              rnorm(50, 1.5, 1),
              rnorm(90, 0, 1),
              rnorm(120, -0.8, 1))
qplot(1:length(sim_mean), sim_mean, geom = "path")

# using changepoint package to identify the CP

##1.  based on mean to detect changepoint
###  binary segmentation method using BIC information criteria maximum of five breaks.
m_binseg <- cpt.mean(sim_mean, penalty = "BIC", method = "BinSeg", Q = 5)
plot(m_binseg, type = "l", xlab = "Index", cpt.width = 4)
cpts(m_binseg) # the CP

## segment neightborhood method using BIC 
m_segneigh <- cpt.mean(sim_mean, penalty = "BIC", method = "SegNeigh", Q = 5)
plot(m_segneigh, type = "l", xlab = "Index", cpt.width = 4)
cpts(m_segneigh) # the CP

## PELT
m_pelt <- cpt.mean(sim_mean, penalty = "BIC", method = "PELT")
plot(m_pelt, type = "l", xlab = "Index", cpt.width = 4)
cpts(m_pelt) # the CP

# *And should we wish, we can also impose a manual penalty with the aid of the following commands:
m_pm <- cpt.mean(sim_mean, penalty = "Manual", pen.value = "1.5 * log(n)", # provide our own penalty function
                      method = "PELT")
plot(m_pm, type = "l", cpt.col = "blue", xlab = "Index", cpt.width = 4)
cpts(m_pm)
```

```{r}
# emperical data
rgdp <- fred("GDPC1", "1947-01-01", "2021-04-01")[, -1] %>% 
    mutate(growth = 100*(price/lag(price) - 1)) %>% 
    drop_na()

m_gdp <- rgdp  %>% 
    pull(growth) %>% 
    cpt.mean(., penalty = "SIC", method = "PELT")
plot(m_gdp)
cpts(m_gdp) # use this method directly we do not know the exact date of previous time
cpt <- rgdp %>% 
    slice(cpts(m_gdp))
cpt
```

##### Change point in variance

```{r}
sim_var <- c(rnorm(100, 0, 1),
              rnorm(50, 0, 2),
              rnorm(90, 0, 1),
              rnorm(120, 0, 0.5))
qplot(1:length(sim_var), sim_var, geom = "path")


##2. based on variance to to detect change point
v_pelt <- cpt.var(sim_var, method = "PELT")
plot(v_pelt, type = "l", cpt.col = "blue", xlab = "Index", cpt.width = 4)
cpts(v_pelt)
```


```{r}
# using gdp growth rate 
v_gdp <- rgdp %>% 
    pull(growth) %>% 
    cpt.var(., method = "PELT")
plot(v_gdp, type = "l", cpt.col = "red", cpt.width = 6)
rgdp %>% 
    slice(cpts(v_gdp))
```

##### Change points in the mean and variance

```{r}
sim_mv <- c(rnorm(100, 0, 1),
             rnorm(50, 1, 2),
             rnorm(90, 0, 1),
             rnorm(120, -0.8, 0.5))

plot.ts(sim_mv)

mv_pelt <- cpt.meanvar(sim_mv, method = "PELT")
plot(mv_pelt)
cpts(mv_pelt)
```


```{r}
mv_gdp <- rgdp %>% 
    pull(growth) %>% 
    cpt.meanvar(., method = "PELT")
plot(mv_gdp, type = "l", cpt.col = "blue", xlab = "Index", cpt.width = 4)
rgdp %>% 
    slice(cpts(mv_gdp))
```



### 1.2  Structural break tests

Problem: Term structural break is synonymous with change point, but would usually refer to cases where there is a change in the regression coefficients

e.g.

$$y_t =  x_t^{\top} \beta_j + \varepsilon_t \;\;\; j = \{ 1, \ldots, m+1 \}$$
The hypothesis that the regression coefficients are constant may be constructed as:

$$\begin{aligned}
H_0 \; : \; \beta_j = \beta_0\\
H_1 \; : \; \beta_j \neq \beta_0
\end{aligned}$$

This test is based on the residual, $\varepsilon_t$, constructing F-statistics:

$$F_i = \frac{\hat{u}^\top\hat{u} - \hat{u}(\tau)^\top \hat{u}(\tau)}{\hat{u}(\tau)^\top \hat{u}(\tau) / \left(T-2k \right)}$$

```{r}
rm(list = ls())
set.seed(123)

x1 <- arima.sim(model = list(ar = 0.9), n = 100)
x2 <- arima.sim(model = list(ma = 0.1), n = 100)
x3 <- arima.sim(model = list(ar = 0.5, ma = 0.3), n = 100)

y <- c((1 + x1),
       x2,
       (0.5 - x3))
plot.ts(y)

## strucchange to test structural change

### Quandt likelihood ratio test

dat <- tibble(ylag0 = y,
              ylag1 = lag(y)
              ) %>%
  drop_na() # because we want to test AR(1) model so only need to generate one lag
qlr <- Fstats(ylag0 ~ ylag1, data = dat) # generate the Quandt Likelihood Ratio (QLR) statistic
breakpoints(qlr)

sctest(qlr, type = "supF") # the significance values for structural break at this point is

plot(qlr)

### CUSUM test

cusum <- efp(ylag0 ~ ylag1, type = "OLS-CUSUM", data = dat)
plot(cusum) # this plot means the data does not contain a structural break.

# this result is incorrect, as we did not define properly about structural break
```


```{r}
rgdp_dat <- fred("GDPC1", "1947-01-01", "2021-04-01")[, -1] %>% 
    mutate(growth = 100*(price/lag(price) - 1),
           growth_lag = lag(growth)) %>% 
    drop_na()
sa_qlr <- Fstats(growth ~ growth_lag, data = rgdp_dat)
breakpoints(sa_qlr)
sctest(sa_qlr, type = "supF")
plot(sa_qlr)
rgdp_dat %>% 
    slice(sa_qlr$breakpoint)
```
Similarly, we can perform the Bai and Perron (2003) test for multiple structural breaks with the aid of the following commands:

```{r}
sa_bp <- breakpoints(growth ~ growth_lag, data = rgdp_dat, breaks = 5)
summary(sa_bp)  # where the breakpoints are

plot(sa_bp, breaks = 5)

rgdp_dat %>% slice(sa_bp$breakpoints)

sa_cusum <- efp(growth ~ growth_lag, data = rgdp_dat, type = "OLS-CUSUM")
plot(sa_cusum)
```



