---
title: "TSRethinking10 Structural vector autoregressive models Emperical study"
author: "Wayne Zhang"
date: "10/9/2021"
output:
  html_document: 
    toc: yes
    toc_depth: 4
    number_sections: true
---
\newcommand{\df}[1]{\textbf{Define #1}}
\newcommand{\remark}{\textbf{Remark}}

The structural form of SVAR will be:

$$y_t = ay_t + A_1^*y_{t - 1} + \cdots + A_p^*y_{t - p} + u_t$$

And rewrite this we have

$$(1 - a)y_t = A_1^*y_{t - 1} + \cdots + A_p^*y_{t - p} + u_t$$

$$y_t = (1 - a)^{-1}A_1^*y_{t - 1} + \cdots + (1 - a)^{-1} A_p^*y_{t - p} + (1 - a)^{-1}u_t$$

def $A_i := (1 - A)^{-1}A_i^*$

We have :

$$y_t = A_1y_{t - 1} + \cdots + A_p y_{t - p} + \varepsilon_t$$

$$(I_K - A_1 L - A_2L^2 - \cdots- A_pL^p)y_t = \varepsilon_t$$


Here, we cannot estimate reduced-form VAR and recover to structural model, so we need to specify structural matrix 

$$A = 
\begin{bmatrix}
1 & 0 & \cdots & 0 \\
a_{21} & 1 &  & 0 \\
\vdots & &  \ddots & \vdots  \\
a_{K1} & a_{K2} & \cdots & 1 \\
\end{bmatrix}$$

**Remark** `A` should at least contains $\frac{K(K - 1)}{2}$ of 0.

$$A(I_K - A_1 L - A_2L^2 - \cdots- A_pL^p)y_t = A\varepsilon_t$$

By Cholesky decomposition(because this matrix is PD), we have 

$$Var(\varepsilon_t) = \Sigma_\varepsilon = U'U$$

$U$ is upper trianguler matrix, then define: $e_t := (U')^{-1}\varepsilon_t; B:= AU'$

we have $e_t \sim N(0, I_K)$

The short-run SVAR model without exogenous variables can be written as:

$$A(I_K - A_1 L - A_2L^2 - \cdots- A_pL^p)y_t = A\varepsilon = Be_t$$

```{r loadingPackage, warning=FALSE, message=FALSE}
library(FinCal)
library(tidyverse)
library(vars)
library(zoo)
```


```{r}
data1 <- stata("use https://www.stata-press.com/data/r17/lutkepohl2\n list qtr", data.out = T) %>% 
    filter(qtr <= 76)
head(data1)

A <- diag(3)
A[lower.tri(A)] <- NA
A
B <- diag(3)
diag(B) <- NA
B
var.est1 <- data.frame(data1$dln_inv, data1$dln_inc, data1$dln_consump) %>% 
    drop_na() %>% 
    VAR(., p = 1, type = "const", season = NULL)
svar_model <- SVAR(var.est1, Amat = A, Bmat = B, max.iter = 10000)

svar_model

A_hat <- svar_model$A
B_hat <- svar_model$B
solve(A_hat) %*% B_hat %*% t(B_hat) %*% t(solve(A_hat))
svar_model$Sigma.U

solve(A_hat) %*% B_hat # U

d("use https://www.stata-press.com/data/r17/lutkepohl2")
d("matrix A = (1,0,0\\ .,1,0\\ .,.,1)")
d("matrix B = (.,0,0\\ 0,.,0\\ 0,0,.)")
d("svar dln_inv dln_inc dln_consump if qtr<=tq(1978q4), aeq(A) beq(B)")
d("matrix Aest = e(A)")
d("matrix Best = e(B)")
d("matrix chol_est = inv(Aest)*Best") #U
d("matrix list chol_est")
d("matrix sig_var = e(Sigma)")
d("matrix list sig_var")
d("matrix chol_var = cholesky(sig_var)")
d("matrix list chol_var", output = T)
```

