---
title: "VECM model"
author: "Weiheng Zhang"
date: '2022-04-25'
output: pdf_document
---

* Goal: To estimate the VEC model, to forecast the price of the gas and compare the predictability of VEC model with those of forecasts using AR/MA/ARMA model.

* Data: 
    * Crude Oil price (FRED/MCOILWTICO);
    * US Regular conventional gas price: (FRED/GASREGCOVM)

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 50), tidy = TRUE)
library(FinMetric)
library(xts)
oil <- fred("MCOILWTICO", "1995-01-01", "2017-03-12") 
gas <- fred("GASREGCOVM", "1995-01-01", "2017-03-12")

loil <- as.xts(oil$MCOILWTICO, order.by = oil$date) %>% 
    log() 
lgas <- as.xts(gas$GASREGCOVM, order.by = gas$date) %>% 
    log()
```

## (a) Create a single time series plot with two log prices logpGASt and logpOILt for the sample 1995M1-2017M4.

```{r}
# (a)
par(mfrow = c(1, 1))
plot(loil, type = "l", main = "WTI vs. Gas", col = "blue", ylim = c(-1, 5))
lines(lgas, col = "red")
```

## (b) Perform unit root tests to determine whether logpGASt and logpOILt are I(0) or I(1).

```{r}
# (b) 
library(urca)
ur.ers(loil, type = "P-test", model = "trend") %>% summary()
ur.ers(lgas, type = "P-test", model = "trend") %>% summary()

# both series could be I(1) or more 

doil <- diff(loil)
dgas <- diff(lgas)

ur.ers(doil, type = "P-test", model = "trend") %>% summary()
ur.ers(dgas, type = "P-test", model = "trend") %>% summary()

# now we cannot reject the null hypothesis that no unit root problem.

par(mfrow=c(1,1))
plot(doil, type='l', main="(log difference)WTI vs. (log difference)Gas", col="blue", ylim=c(-0.4, 0.4))
lines(dgas, col="red")
```

## (c)  Determine the number of lags to include in cointegration analysis using Schwarz information criterion. Run the Johansen’s trace and maximum eigenvalue cointegration tests for ( logpGASt and logpOILt ) using the sample 1995M1-2010M12. Use time series plots from (a) as a guide to determine the specification of the deterministic components in the cointegration test (i.e. whether to use Case 2, Case 3, of Case 4 cointegration test).

```{r}
# (c)
library(vars)
y <- cbind(loil, lgas)
colnames(y) <- c("log.wti", "log.gas")
y <- na.trim(y)

y <- y["1995/2010-12-31"]
# determine number of lags to be included in cointegration test and in VEC model
y_VAR_order <- VARselect(y, type = "const")
y_VAR_order
nlags <- y_VAR_order$selection["SC(n)"]
nlags

ca.jo(y, ecdet = "const", type = "trace", K = nlags, spec = "transitory") %>% 
    summary()
y.CA <- ca.jo(y, ecdet="const", type="eigen", K=nlags, spec="transitory")
summary(y.CA)


lttest(y.CA, r = 1)
```

The result of lttest which tests for the presence of the constant term in the cointegration relationship suggests case 2.

## (d) Based on your results in (c), use the 1995M1-2010M12 sample to estimate a bivariate VEC model for logpGASt and logpOILt.


```{r}
# estimate unrestricted VEC model
y.VEC <- cajorls(y.CA, r = 1)
y.VEC
```

# (e) Are the adjustment parameters $\alpha_1$ and $\alpha_2$ in the estimated VEC model statistically significant? Are their signs consistent with error correction mechanism that moves the system back to the long run equilibrium, whenever there is a disruption and zt−1 is not zero?

```{r}
# to see t-statistics and p-values
summary(y.VEC$rlm)
```


## 

```{r}
#Q.F

# test for restricted adjustment parameters alpha
rest.alpha <- matrix(c(1,0), c(2,1))
y.CA.ralpha <- alrtest(y.CA, A=rest.alpha, r=1)
summary(y.CA.ralpha)

# Q.G

# vec2varX supports restricted VEC models
y.VAR <- vec2var(y.CA, r=1)
# y.VAR.rest <- vec2var(y.CA, A=rest.alpha, r=1)

# forecast using VAR in levels
y.VAR.fcst <- predict(y.VAR, n.ahead=76, ci=0.95)
par( mar=c(4,4,2,1), cex=0.75)
plot(y.VAR.fcst)
```











